Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets

### 2021-07-06 main.tex
Fin première lecture P.

\begin{definition}
\label{def:transducer} \label{def:SWT}
A \emph{symbolic-weighted transducer} (\SWT)
over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$
%the input and output alphabets~$\Sigma$ and $\Delta$ with label theory $\bar\Phi$, and the semiring $\Semiring$ 
is a tuple
$T = \< Q, \init, \bar{\wei}, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$   %\Phi_{\Sigma, \Delta}
(respectively $\mathsf{out} : Q \to \Semiring$)  %\Phi_{\Sigma, \Delta}
are functions defining the weight for entering
(respectively leaving) computation in a state,  
and $\bar{\wei}$ is a triplet of transition functions 
$\wei_{10}: Q \times Q \to \Phi_{\Sigma}$,
$\wei_{01}: Q \times Q \to \Phi_{\Delta}$, and 
$\wei_{11}: Q \times Q \to \Phi_{\Sigma, \Delta}$.
\end{definition}
%
\noindent
We call \emph{number of transitions} of $T$ the number of pairs of states 
$q, q' \in Q$ such that $\wei_{10}$ or $\wei_{01}$ or $\wei_{11}$
is not the constant $\zero$.
%
\noindent
For convenience, we shall sometimes present transitions 
as functions of  
$Q \times (\Sigma \cup \{ \varepsilon \}) \times (\Delta \cup \{ \varepsilon \}) \times Q \to \Semiring$,
overloading the function names,
such that, for all $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$, 
\[
\begin{array}{rcll}
\wei_{10}(q, a, \varepsilon, q') & = & \phi(a) & %\wei_\Sigma(q, q')(a)
\quad\mathrm{where~} \phi = \wei_{10}(q, q') \in \Phi_\Sigma,\\
\wei_{01}(q, \varepsilon, b, q') & = & \psi(b) & 
\quad\mathrm{where~} \psi = \wei_{01}(q, q') \in \Phi_\Delta,\\
\wei_{11}(q, a, b, q') & = & \eta(a, b) & 
\quad\mathrm{where~} \eta = \wei_{11}(q, q') \in \Phi_{\Sigma, \Delta}.\\
\end{array}      
\]
%
\noindent 
% The symbolic-weighted transducer
The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^* \times \Delta^*$,
$s$ and $t$, being respectively called \emph{input} and \emph{output} word.
% \noindent
More precisely, $T$ defines a mapping 
from $\Sigma^* \times \Delta^*$ into~$\Semiring$,
based on an intermediate function $\weight_T$
defined recursively, for every states $q, q' \in Q$, 
and every strings $\< s, t> \in \Sigma^* \times \Delta^* \setminus \{ \< \varepsilon, \varepsilon >\}$ considered as concatenation
of the symbol $a \in \Sigma$ (resp. $b \in \Delta$) with a word $u \in \Sigma^*$ (resp. $v in \Delta^*$), by :
\lydia{added u and v def}

@@

%
We recall that, by convention (Section~\ref{sec:semiring}), 
an empty sum with $\bigoplus$ is equal to~$\zero$. 
%
Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
when reading respectively $a$ and $b$ at the current positions in the input and output words, 
increment the current position in the input word if and only if $i = 1$, 
and in the output word iff $j = 1$, %(otherwise, do not change it),
and change state from $q$ to $q'$.
When $a = \varepsilon$ (resp. $b = \varepsilon$), the current symbol 
in the input (resp. output) is not read.
%
%In contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
%the input and output symbols at current positions are always read by transitions, 
%even when they do not change the reading position the head's position.
%This is an important feature in the case of an infinite alphabet in 
%order to compare input and output symbols.
%which cannot be stored in the finite memory of the transducer.
%
Since~$\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression 
will be ignored in the sum, meaning that there is no possible transition
from state $q$ into state $q'$ while reading $a$ and $b$.
This is analogous to the case of a transition's guard not satisfied by $\<a, b>$ for 
symbolic transducers.

@@

\begin{example}
In Common Western Music Notation~\cite{Gould11Notation}, 
several symbols may be used to represent one single sounding event. 
\lydia{unique $\to$ similar}
\florent{similar $\to$ single}
For instance, several notes can be combined with a tie, 
like in $\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,  
and one note can be augmented by half its duration with a dot like in~\musQuarterDotted{}.
\lydia{modif.}
These notations are perceived equivalent when played, as their duration is equal, yet the notation is different.
We thus want to be able to compare a music score with music played by a performer. 
\lydia{changed end}
%
\noindent 
We propose a small weighted transducer model that calculates the distance bewteen an input sequence of sounding 
events (music "performance") 
to an output sequence of written events (music "score").
% simple pointwise distance between two sequences of timestamped events **
%
Let us consider the tropical (\emph{min-plus}) semiring~$\Semiring$ 
of Figure~\ref{fig:semirings} and 
let $\Sigma = \mathbb{R}_+$ be an input alphabet of event dates
and $\Delta = \{ \mathsf{e}, \mathsf{-} \} \times \mathbb{R}_+$ 
be an output alphabet of symbols with timestamps. 
A symbol $\< \mathsf{e}, d > \in \Delta$ represents an event starting at date $d$, 
and $\< \mathsf{-}, d >$ is a continuation of the previous event.
%More precisely, we let 
%$\Sigma = \{ a, - \} \times \R_+$ and 
%This example of $\Delta$ is motivated by the case of music notation, 
%where several notated events (notes) can be tied together, 
%with a \emph{tie} or a \emph{dot}
%meaning that they will be played as a unique sounding event.
%The timestamp of $a \in \Sigma$, denoted by $\mathsf{t}(a)$, is expressed as a rational number.

We consider a \SWT with two states $q_0$ and $q_1$ whose purpose 
is to compare a recorded performance $s \in \Sigma^*$
with a notated music sheet $t \in \Delta^*$.
One timestamp $d_i \in \Sigma$ may correspond 
to one notated event $\<\mathsf{e}, d'_i> \in \Delta$, in which case 
the weight value computed by the \SWT is the time distance between both
(see transitions $\wei_{11}$ below).
%
If $\<\mathsf{e}, d'_i>$ is followed by continuations 
$\<\mathsf{-}, d'_{i+1}>$..., they are just skipped with no cost (transitions $\wei_{01}$ or weight $\one$).
%transitions
\[
\begin{array}{rclcrcl}
\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
\wei_{01}(q_0, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one & &
\wei_{01}(q_1, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one\\
\wei_{10}(q_0, d, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
\end{array}
\]
%
We also must able to take performing errors into account, while still being able to compare with the score, 
since a performer could, for example, play an unwritten extra note. 
\lydia{reformulated this sentence} 
%
This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$, 
switching from state $q_0$ (normal) to $q_1$ (error).
The transitions in the second column below switch back to the normal state $q_0$.
% the metric computed by the \SWT is the smallest sum of point wise distances 
% between dates of input and output events.
At last, we let $q_0$ be the only initial and final state, with
$\init(q_0) = \final(q_0) = \one$, and 
$\init(q_1) = \final(q_1) = \zero$.

@@

\noindent
The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
are defined similarly as the transducers of Definition~\ref{def:SWT}, 
by simply omitting the output symbols.
%
%In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
%
\begin{definition} \label{def:SWA}
A \emph{symbolic-weighted automaton} (\SWA)
over $\Sigma$, $\Semiring$ and $\bar\Phi$
%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$ 
is a tuple
$A = \< Q, \init, {\wei_1}, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$ %\Phi_\Sigma$, 
(respectively $\mathsf{out} : Q \to \Semiring$)  %\Phi_\Sigma,$
are functions defining the weight for entering
(respectively leaving) computation in a state, 
and ${\wei_1}$ is a transition functions 
from $Q \times Q$ into~$\Phi_{\Sigma}$.
\end{definition}
%      
\noindent
As above in the case of \SWT, 
when $\wei_1(q, q') = \phi \in \Phi_\Sigma$, 
%respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
we may write $\wei_1(q, a, q')$ for~$\phi(a)$. 
%respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
%$\wei_1: Q \times \Sigma \times Q \to \Semiring$, 
The computation of $A$ on words $s \in \Sigma^*$
is defined with an intermediate function $\weight_A$, 
defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$,
%
\begin{align}
\weight_A(q, \varepsilon, q) & = \one\\ %\final(q)\\
\weight_A(q, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
\weight_A(q, au, q') & =  \displaystyle\bigoplus_{q'' \in Q} 
    \wei_{1}(q, a, q'') \otimes \weight_A(q'', u, q')\nonumber
\label{eq:SWA-weight}
\end{align}
%
\noindent
and the weight value associated by $A$ to 
$s \in \Sigma^*$ is defined as follows: %$s = s_1\ldots s_n \in \Sigma^+$ 
\begin{equation}
A(s)  = 
\displaystyle\bigoplus_{q, q' \in Q} \init(q) 
\mathop{\otimes} \weight_A(q, s, q') \mathop{\otimes} \final(q') 
\label{eq:SWA-value}
\end{equation}

@@

\begin{proposition} \label{prop:epsilon}
Given a \SWT $T$ over $\Sigma$, $\Delta$, 
$\Semiring$ commutative, bounded and complete,
and $\bar\Phi$ effective,
and a $\SWA$ $A$ over $\Sigma$, $\Semiring$ and $\bar\Phi$,
there exists an effectively constructible \SWA 
$B_{T, A}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
such that for all $t \in \Delta^*$, 
$B_{T, A}(t) = \displaystyle\bigoplus_{s\in \Sigma^*} A(s) \otimes T(s, t)$.
\end{proposition}
%
\begin{proof}
Let $T = \< Q, \init_T, \bar{\wei}, \final_T >$,
where $\bar{\wei}$ contains $\wei_{10}$, $\wei_{01}$, and $\wei_{11}$,
from $Q \times Q$ into respectively 
$\Phi_{\Sigma}$, $\Phi_{\Delta}$, and $\Phi_{\Sigma, \Delta}$,
and let $A = \< P, \init_A, \wei_1, \final_A >$
with $\wei_1: Q \times Q \to \Phi_{\Sigma}$.
%
\noindent
The state set of $B_{T, A}$ will be $Q' = P \times Q$.
The entering, leaving and transition functions of $B_{T, A}$ will 
simulate synchronized computations of $A$ and $T$, 
while reading an output word of $\Delta^*$.
%
Its state entering functions is defined 
for all $p \in P$, $q \in Q$ 
by $\init'(p, q) = \init_A(p) \otimes \init_T(q)$.
The transition function $\wei'_1$ will roughly perform 
a synchronized product of transitions defined by $\wei_1$, 
$\wei_{01}$ ($T$ reading in output word and not in input word)
and $\wei_{11}$ ($T$ reading in output word and input word).
%
Moreover, $\wei'_1$ also needs to simulate transitions 
defined by $\wei_{10}$: $T$ reading in input word and not in output word.
Since $B_{T, A}$  will read only in the output word, such a transition corresponds
to an $\varepsilon$-transition of $\SWA$, 
but $\SWA$ have been defined without $\varepsilon$-transitions.
Therefore, in order to take care of this case, we perform an on-the-fly
suppression of $\varepsilon$-transition in the $\SWA$ in construction, 
following the algorithm of~\cite{LombardySakarovitch12ciaa}. 
%
% Initialize state leaving functions i
%and $\final'(p, q) = \final_A(p) \otimes \final_T(q)$.

@@

%\subsection{Definition}
\noindent
\label{sec:SWVPA-def}
Let $\Omega$ be a countable alphabet 
%finite (large) or infinite,
that we assume partitioned into three 
subsets~$\Omegai$, $\Omegac$, $\Omegar$,
whose elements are respectively called 
\emph{internal}, \emph{call} and \emph{return} symbols.
% names are  coined by application to functional program verification
% \begin{itemize}
% \item a set $\Omegai$ of \emph{internal symbols} denoted $a$,
% \item a set $\Omegac$ of \emph{call symbols} denoted $\call{a}$,
% \item a set $\Omegar$ of \emph{return symbols} denoted $\return{a}$.
% \end{itemize}
Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative and complete semiring and let  
$\bar\Phi = \< \Phii, \Phic, \Phir, \Phici, \Phicc, \Phicr>$ 
be a label theory over $\Semiring$
%In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
%we shall write respectively 
where $\Phii$, $\Phic$, $\Phir$ and~$\Phicx$ (with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$) 
stand respectively 
for~$\Phi_\Omegai$, $\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omega_\mathsf{x}}$.
%
%Moreover, we extend this theory with a set $\Phii$ 
%of unary functions in $\Omegai \to \Semiring$,
%closed under $\oplus$ and $\otimes$.

>>>>

\begin{definition}
\label{def:transducer} \label{def:SWT}
A \emph{symbolic-weighted transducer} (\SWT)
over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$
%the input and output alphabets~$\Sigma$ and $\Delta$ with label theory $\bar\Phi$, and the semiring $\Semiring$ 
is a tuple
$T = \< Q, \init, \bar{\wei}, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$   %\Phi_{\Sigma, \Delta}
(respectively $\mathsf{out} : Q \to \Semiring$)  %\Phi_{\Sigma, \Delta}
are functions defining the weight for entering
(respectively leaving) computation in a state,  
and $\bar{\wei}$ is a triplet of transition functions 
$\wei_{10}: Q \times Q \to \Phi_{\Sigma}$,
$\wei_{01}: Q \times Q \to \Phi_{\Delta}$, and 
$\wei_{11}: Q \times Q \to \Phi_{\Sigma, \Delta}$.
\end{definition}
%
\noindent
We call \emph{number of transitions} of $T$ the number of pairs of states 
$q, q' \in Q$ such that $\wei_{10}$ or $\wei_{01}$ or $\wei_{11}$
is not the constant $\zero$.
%
\noindent
For convenience, we shall sometimes present transitions 
as functions of  
\philippe{I missed sth: what is this $\varepsilon$? Intuitively clear but not defined?}
$Q \times (\Sigma \cup \{ \varepsilon \}) \times (\Delta \cup \{ \varepsilon \}) \times Q \to \Semiring$,
overloading the function names,
such that, for all $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$,

\[
\begin{array}{rcll}
\wei_{10}(q, a, \varepsilon, q') & = & \phi(a) & %\wei_\Sigma(q, q')(a)
\quad\mathrm{where~} \phi = \wei_{10}(q, q') \in \Phi_\Sigma,\\
\wei_{01}(q, \varepsilon, b, q') & = & \psi(b) & 
\quad\mathrm{where~} \psi = \wei_{01}(q, q') \in \Phi_\Delta,\\
\wei_{11}(q, a, b, q') & = & \eta(a, b) & 
\quad\mathrm{where~} \eta = \wei_{11}(q, q') \in \Phi_{\Sigma, \Delta}.\\
\end{array}      
\]
%
\noindent 
% The symbolic-weighted transducer
The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^* \times \Delta^*$,
$s$ and $t$, being respectively called \emph{input} and \emph{output} word.
% \noindent
More precisely, $T$ defines a mapping 
from $\Sigma^* \times \Delta^*$ into~$\Semiring$,
based on an intermediate function $\weight_T$
defined recursively, for every states $q, q' \in Q$, 
and every pairs of strings $\< s, t> \in \Sigma^* \times \Delta^* \setminus \{ \< \varepsilon, \varepsilon >\}$,
each  considered as the concatenation
of the symbol $a \in \Sigma$ (resp. $b \in \Delta$) with a word $u \in \Sigma^*$ (resp. $v \in \Delta^*$), by :
\lydia{added u and v def}

@@

%
We recall that, by convention (Section~\ref{sec:semiring}), 
an empty sum with $\bigoplus$ is equal to~$\zero$. 
%
\philippe{OK tout ça se lit bien :-)}
Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
when reading respectively $a$ and $b$ at the current positions in the input and output words, 
increment the current position in the input word if and only if $i = 1$, 
and in the output word iff $j = 1$, %(otherwise, do not change it),
and change state from $q$ to $q'$.
When $a = \varepsilon$ (resp. $b = \varepsilon$), the current symbol 
in the input (resp. output) is not read.
%
%In contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
%the input and output symbols at current positions are always read by transitions, 
%even when they do not change the reading position the head's position.
%This is an important feature in the case of an infinite alphabet in 
%order to compare input and output symbols.
%which cannot be stored in the finite memory of the transducer.
%
Since~$\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression 
will be ignored in the sum, meaning that there is no possible transition
from state $q$ into state $q'$ while reading $a$ and $b$.
This is analogous to the case of a transition's guard not satisfied by $\<a, b>$ for 
symbolic transducers.

@@

\philippe{Je crois qu'il faudrait numéroter les exemples indépendamment des définitions. Cet exemple est le premier
qui donne des détails sur l'application visée. Il arrive peut-être un peu tard et est long. On pourrait introduire
la motivation dans l'intro, et développer des petits exemples au fur et à mesure.}
\begin{example}
In Common Western Music Notation~\cite{Gould11Notation}, 
several symbols may be used to represent one single sounding event. 
\lydia{unique $\to$ similar}
\florent{similar $\to$ single}
For instance, several notes can be combined with a tie, 
like in $\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,  
and one note can be augmented by half its duration with a dot like in~\musQuarterDotted{}.
\lydia{modif.}
These notations are perceived equivalent when played, as their duration is equal, yet the notation is different.
We thus want to be able to compare a music score with music played by a performer. 
\lydia{changed end}
%
\noindent 
We propose a small weighted transducer model that calculates the distance bewteen an input sequence of sounding 
events (music "performance") 
to an output sequence of written events (music "score").
% simple pointwise distance between two sequences of timestamped events **
%
Let us consider the tropical (\emph{min-plus}) semiring~$\Semiring$ 
of Figure~\ref{fig:semirings} and 
let $\Sigma = \mathbb{R}_+$ be an input alphabet of event dates
and $\Delta = \{ \mathsf{e}, \mathsf{-} \} \times \mathbb{R}_+$ 
be an output alphabet of symbols with timestamps. 
A symbol $\< \mathsf{e}, d > \in \Delta$ represents an event starting at date $d$, 
and $\< \mathsf{-}, d >$ is a continuation of the previous event.
%More precisely, we let 
%$\Sigma = \{ a, - \} \times \R_+$ and 
%This example of $\Delta$ is motivated by the case of music notation, 
%where several notated events (notes) can be tied together, 
%with a \emph{tie} or a \emph{dot}
%meaning that they will be played as a unique sounding event.
%The timestamp of $a \in \Sigma$, denoted by $\mathsf{t}(a)$, is expressed as a rational number.

We consider a \SWT with two states $q_0$ and $q_1$ whose purpose 
is to compare a recorded performance $s \in \Sigma^*$
with a notated music sheet $t \in \Delta^*$.
One timestamp $d_i \in \Sigma$ may correspond 
to one notated event $\<\mathsf{e}, d'_i> \in \Delta$, in which case 
the weight value computed by the \SWT is the time distance between both
(see transitions $\wei_{11}$ below).
%
If $\<\mathsf{e}, d'_i>$ is followed by continuations 
$\<\mathsf{-}, d'_{i+1}>$..., they are just skipped with no cost (transitions $\wei_{01}$ or weight $\one$).
%transitions
\[
\begin{array}{rclcrcl}
\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
\wei_{01}(q_0, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one & &
\wei_{01}(q_1, \varepsilon, \< \mathsf{-}, d'>, q_0) & = & \one\\
\wei_{10}(q_0, d, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
\end{array}
\]
%
We also must be able to take performing errors into account, while still being able to compare with the score, 
since a performer could, for example, play an unwritten extra note. 
\lydia{reformulated this sentence} 
%
This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$, 
switching from state $q_0$ (normal) to $q_1$ (error).
The transitions in the second column below switch back to the normal state $q_0$.
% the metric computed by the \SWT is the smallest sum of point wise distances 
% between dates of input and output events.
At last, we let $q_0$ be the only initial and final state, with
$\init(q_0) = \final(q_0) = \one$, and 
$\init(q_1) = \final(q_1) = \zero$.

@@

\noindent
The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
are defined similarly as the transducers of Definition~\ref{def:SWT}, 
by simply omitting the output symbols.
%
%In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
%
\begin{definition} \label{def:SWA}
A \emph{symbolic-weighted automaton} (\SWA)
over $\Sigma$, $\Semiring$ and $\bar\Phi$
%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$ 
is a tuple
$A = \< Q, \init, {\wei_1}, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Semiring$ %\Phi_\Sigma$, 
(respectively $\mathsf{out} : Q \to \Semiring$)  %\Phi_\Sigma,$
are functions defining the weight for entering
(respectively leaving) computation in a state, 
and ${\wei_1}$ is a transition function 
from $Q \times Q$ into~$\Phi_{\Sigma}$.
\end{definition}
%      
\noindent
As above in the case of \SWT, 
when $\wei_1(q, q') = \phi \in \Phi_\Sigma$, 
%respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
we may write $\wei_1(q, a, q')$ for~$\phi(a)$. 
%respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
%$\wei_1: Q \times \Sigma \times Q \to \Semiring$, 
The computation of $A$ on words $s \in \Sigma^*$
is defined with an intermediate function $\weight_A$, 
defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$,
%
\begin{align}
\weight_A(q, \varepsilon, q) & = \one\\ %\final(q)\\
\weight_A(q, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
\weight_A(q, au, q') & =  \displaystyle\bigoplus_{q'' \in Q} 
    \wei_{1}(q, a, q'') \otimes \weight_A(q'', u, q')\nonumber
\label{eq:SWA-weight}
\end{align}
%
\noindent
and the weight value associated by $A$ to 
$s \in \Sigma^*$ is defined as follows: %$s = s_1\ldots s_n \in \Sigma^+$ 
\begin{equation}
A(s)  = 
\displaystyle\bigoplus_{q, q' \in Q} \init(q) 
\mathop{\otimes} \weight_A(q, s, q') \mathop{\otimes} \final(q') 
\label{eq:SWA-value}
\end{equation}

@@

\begin{proposition} \label{prop:epsilon}
Given a \SWT $T$ over $\Sigma$, $\Delta$, 
$\Semiring$ commutative, bounded and complete,
and $\bar\Phi$ effective,
and a $\SWA$ $A$ over $\Sigma$, $\Semiring$ and $\bar\Phi$,
there exists an effectively constructible \SWA 
$B_{T, A}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
such that for all $t \in \Delta^*$, 
$B_{T, A}(t) = \displaystyle\bigoplus_{s\in \Sigma^*} A(s) \otimes T(s, t)$.
\end{proposition}
%
\begin{proof}
Let $T = \< Q, \init_T, \bar{\wei}, \final_T >$,
where $\bar{\wei}$ contains $\wei_{10}$, $\wei_{01}$, and $\wei_{11}$,
from $Q \times Q$ into respectively 
$\Phi_{\Sigma}$, $\Phi_{\Delta}$, and $\Phi_{\Sigma, \Delta}$,
and let $A = \< P, \init_A, \wei_1, \final_A >$
with $\wei_1: Q \times Q \to \Phi_{\Sigma}$.
%
\noindent
The state set of $B_{T, A}$ will be $Q' = P \times Q$.
The entering, leaving and transition functions of $B_{T, A}$ will 
simulate synchronized computations of $A$ and $T$, 
while reading an output word of $\Delta^*$.
%
Its state entering functions is defined 
for all $p \in P$, $q \in Q$ 
by $\init'(p, q) = \init_A(p) \otimes \init_T(q)$.
The transition function $\wei'_1$ will roughly perform 
a synchronized product of transitions defined by $\wei_1$, 
$\wei_{01}$ ($T$ reading in output word and not an input word)
and $\wei_{11}$ ($T$ reading both an input word and an output word).
%
Moreover, $\wei'_1$ also needs to simulate transitions 
defined by $\wei_{10}$: $T$ reading in input word and not an output word.
Since $B_{T, A}$  will read only in the output word, such a transition corresponds
to an $\varepsilon$-transition of $\SWA$, 
but $\SWA$ have been defined without $\varepsilon$-transitions.
Therefore, in order to take care of this case, we perform an on-the-fly
suppression of $\varepsilon$-transition in the $\SWA$ in construction, 
following the algorithm of~\cite{LombardySakarovitch12ciaa}. 
%
% Initialize state leaving functions i
%and $\final'(p, q) = \final_A(p) \otimes \final_T(q)$.

@@

%\subsection{Definition}
\noindent
\label{sec:SWVPA-def}
Let $\Omega$ be a countable alphabet 
%finite (large) or infinite,
that we assume partitioned into three 
subsets~$\Omegai$, $\Omegac$, $\Omegar$,
whose elements are respectively called 
\emph{internal}, \emph{call} and \emph{return} symbols.
\philippe{Là je crois qu'il faudrait expliquer ces Omega, je commence à fatiguer et je suis un
peu largué par toutes ces définitions. J'intuite qu'il s'agit des symboles, parenthèses ouvrantes
et fermantes? Pourquoi il faut un alphabet pour les parenthèses?}
% names are  coined by application to functional program verification
% \begin{itemize}
% \item a set $\Omegai$ of \emph{internal symbols} denoted $a$,
% \item a set $\Omegac$ of \emph{call symbols} denoted $\call{a}$,
% \item a set $\Omegar$ of \emph{return symbols} denoted $\return{a}$.
% \end{itemize}
Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative and complete semiring and let  
$\bar\Phi = \< \Phii, \Phic, \Phir, \Phici, \Phicc, \Phicr>$ 
be a label theory over $\Semiring$
%In order to simplify notations, %and following the definition of Section~\ref{section:transducer}, 
%we shall write respectively 
where $\Phii$, $\Phic$, $\Phir$ and~$\Phicx$ (with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$) 
stand respectively 
for~$\Phi_\Omegai$, $\Phi_\Omegac$, $\Phi_\Omegar$ and~$\Phi_{\Omegac, \Omega_\mathsf{x}}$.
%
%Moreover, we extend this theory with a set $\Phii$ 
%of unary functions in $\Omegai \to \Semiring$,
%closed under $\oplus$ and $\otimes$.

\begin{definition}
A \emph{Symbolic Weighted Visibly Pushdown Automata} (\SWVPA) 
over  $\Omega = \Omegai \uplus \Omegac \uplus \Omegar$, $\Semiring$ and $\bar\Phi$ 
is a tuple $A = \< Q, P, \init, \bar\wei, \final >$,
where $Q$ is a finite set of states, 
\philippe{Est-ce que tout le monde sait ce qu'est un pushdown automata? Je suppose que c'est
lié à la pile.}
$P$ is a finite set of stack symbols, 
$\mathsf{in} : Q \to \Semiring$ 
(respectively $\mathsf{out} : Q \to \Semiring$)
are functions defining the weight for entering
(respectively leaving) a state, 
and $\bar\wei$ is a sextuplet composed of the transition functions :
$\weii : Q \times P \times Q \to \Phici$,  
$\weiei : Q \times Q \to \Phii$,  
$\weic : Q \times P \times Q \times P \to \Phicc$,  
$\weiec : Q \times P \times Q \to \Phic$,  
$\weir : Q \times P \times Q \to \Phicr$,  
$\weier : Q \times Q \to \Phir$.
%and 
%$\weiex : Q \times Q \to \Phix$ 
%with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$.
\end{definition}
%
Similarly as in Section~\ref{section:transducer}, 
we extend the above transition functions as follows
for all $q, q' \in Q$, $p \in P$, 
$a \in \Omegai$, 
$\call{c} \in \Omegac$, 
$\return{r} \in \Omegar$, 
overloading their names: % for simplicity:
\[
\begin{array}{lll}
\weii: Q \times \Omegac \times P \times \Omegai \times Q \to \Semiring & 
\weii(q, c, p, a, q') = \eta_\mathsf{ci}(c, a) & 
\mathrm{where~} \eta_\mathsf{ci} = \weii(q, p, q'),\\
%
\weiei: Q \times \Omegai \times Q \to \Semiring & 
\weiei(q, a, q') = \phi_\mathsf{i}(a) &
\mathrm{where~} \phi_\mathsf{i} = \weiei(q, q').\\
%
\weic: Q \times \Omegac \times P \times  \Omegac \times P \times Q \to \Semiring & 
\weic(q, c, p, \call{c'}, p', q') = \eta_\mathsf{cc}(c, \call{c'}) & 
\mathrm{where~} \eta_\mathsf{cc} = \weic(q, p, p', q'),\\
%
\weiec: Q \times \Omegac \times P \times Q \to \Semiring & 
\weiec(q, {\call{c}}, p, q') = \phi_\mathsf{c}({\call{c}}) &
\mathrm{where~} \phi_\mathsf{c} = \weiec(q, p, q').\\
%
\weir: Q \times \Omegac \times P \times \Omegar \times Q \to \Semiring & 
\weir(q, {\call{c}},  p, {\return{r}}, q') = \eta_\mathsf{cr}({\call{c}},  {\return{r}}) & 
\mathrm{where~} \eta_\mathsf{cr} = \weir(q, p, q'),\\
%
\weier: Q \times \Omegar \times Q \to \Semiring & 
\weier(q, {\return{r}}, q') = \phi_\mathsf{r}({\return{r}}) &
\mathrm{where~} \phi_\mathsf{r} = \weier(q, q').\\
\end{array}      
\]
